---
title: "Chapter 5 Practice"
author: "Keith Williams"
date: "1/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(tidyverse)
ggplot2::theme_set(theme_minimal())
```

# Notes

```{r}
# number of plants
N <- 100

# simulate initial heights
h0 <- rnorm(N, 10, 2)

# assign treatments and simulate fungus and growth
treatment <- rep(0:1 , each=N/2)
fungus <- rbinom(N , size=1 , prob=0.5 - treatment*0.4)
h1 <- h0 + rnorm(N, 5 - 3*fungus)

# compose a clean data frame
dat <- tibble(h0=h0, h1=h1, treatment=treatment, fungus=fungus)

ggplot(dat, aes(h0, h1, color=as.factor(fungus))) + 
  geom_point()

ggplot(dat, aes(h0, h1, color=as.factor(treatment))) +
  geom_point()

summary_fungus <- dat %>% 
  group_by(treatment) %>% 
  summarise(n = n(), f = sum(fungus))
summary_fungus

# test effect of treatment on fungus
prop.test(x = summary_fungus$f, n = summary_fungus$n)

## R code 5.42
m5.13 <- quap(
    alist(
        h1 ~ dnorm(mu, sigma),
        mu <- a + bh*h0 + bt*treatment + bf*fungus,
        a ~ dnorm(0, 100),
        c(bh, bt, bf) ~ dnorm(0, 10),
        sigma ~ dunif(0, 10)
    ),
    data=dat)
precis(m5.13)

## R code 5.43
m5.14 <- quap(
    alist(
        h1 ~ dnorm(mu, sigma),
        mu <- a + bh*h0 + bt*treatment,
        a ~ dnorm(0, 100),
        c(bh, bt) ~ dnorm(0, 10),
        sigma ~ dunif(0, 10)
    ),
    data=dat)
precis(m5.14)
```

```{r}
## R code 5.44
data(Howell1)
glimpse(Howell1)

## R code 5.45
m5.15 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + bm*male ,
        a ~ dnorm( 178 , 100 ) ,
        bm ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=Howell1)
precis(m5.15)

## R code 5.46
post <- extract.samples(m5.15) %>% 
  mutate(mu_male = post$a + post$bm)

post %>% 
  select(female = a, male = mu_male) %>% 
  gather("sex", "avg_height", female, male) %>% 
  ggplot(aes(avg_height, fill = sex)) + 
  geom_density(alpha = 0.5)

PI(post$mu_male)

## reparamaterized
m5.15b <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- af*(1-male) + am*male ,
        af ~ dnorm( 178 , 100 ) ,
        am ~ dnorm( 178 , 100 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=Howell1  
)

post <- extract.samples(m5.15b)
post %>% 
  select(female = af, male = am) %>% 
  gather("sex", "avg_height") %>% 
  ggplot(aes(avg_height, fill = sex)) + 
  geom_density(alpha = 0.5)

```

```{r}
data(milk)
glimpse(milk)
milk <- mutate(milk, clade_id = coerce_index(clade))
select(milk, clade, clade_id)

ggplot(milk, aes(clade, kcal.per.g)) + 
  geom_point(alpha = 0.7, size = 3)

m5.16_alt <- quap(
  flist = alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a[clade_id],
    a[clade_id] ~ dnorm(0.6, 10),
    sigma ~ dunif(0, 10)
  ),
  data = milk
)
precis(m5.16_alt, depth = 2)
```

# Practice  

## Hard

### 5H1  

```{r}
data(foxes)
glimpse(foxes)

plot_area <- ggplot(foxes, aes(x = area)) +
  geom_point(aes(y = weight, color = as.factor(group)))

plot_groupsize <- ggplot(foxes, aes(x = groupsize)) + 
  geom_point(aes(y = weight, color = as.factor(group)))

m_area <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta * area,
    alpha ~ dnorm(4, 10),
    beta ~ dnorm(0, 4),
    sigma ~ dexp(2)
  ),
  data = foxes
)
p_area <- extract.samples(m_area)
h_area <- tibble(
  area = seq(min(foxes$area), max(foxes$area), length.out = nrow(foxes))
) %>% 
  mutate(
    weight_hat = purrr::map(area, ~p_area$alpha + .x * p_area$beta),
    map_weight = purrr::map_dbl(weight_hat, mean),
    hpdi = purrr::map(weight_hat, ~HPDI(.x, 0.95)),
    lower95 = purrr::map_dbl(hpdi, ~.x[1]),
    upper95 = purrr::map_dbl(hpdi, ~.x[2])
  )

plot_area +
  geom_line(data = h_area, aes(y = map_weight)) +
  geom_ribbon(data = h_area, aes(ymin = lower95, ymax = upper95), alpha = 0.1)

m_groupsize <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta * groupsize,
    alpha ~ dnorm(4, 10),
    beta ~ dnorm(0, 4),
    sigma ~ dexp(2)
  ),
  data = foxes
)

p_groupsize <- extract.samples(m_groupsize)
h_groupsize <- tibble(
  groupsize = seq(min(foxes$groupsize), max(foxes$groupsize), length.out = nrow(foxes))
) %>% 
  mutate(
    weight_hat = purrr::map(groupsize, ~p_groupsize$alpha + .x * p_groupsize$beta),
    map_weight = purrr::map_dbl(weight_hat, mean),
    hpdi = purrr::map(weight_hat, ~HPDI(.x, 0.95)),
    lower95 = purrr::map_dbl(hpdi, ~.x[1]),
    upper95 = purrr::map_dbl(hpdi, ~.x[2])
  )

plot_groupsize +
  geom_line(data = h_groupsize, aes(y = map_weight)) +
  geom_ribbon(data = h_groupsize, aes(ymin = lower95, ymax = upper95), alpha = 0.1)
```

Both appear to have no value in predicting weight on their own. Maybe groupsize has a small, negative correlation.  

### 5H2  

```{r}
m_h2 <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta1 * area + beta2 * groupsize,
    alpha ~ dnorm(4, 10),
    beta1 ~ dnorm(0, 5),
    beta2 ~ dnorm(0, 5),
    sigma ~ dexp(1)
  ),
  data = foxes
)
precis(m_h2)
p_foxes <- extract.samples(m_h2, 1000)

p_foxes %>% 
  mutate(intercept = alpha + beta2 * mean(foxes$groupsize)) %>% 
ggplot(aes(x = area, y = weight)
) +
  geom_abline(aes(intercept = intercept, slope = beta1), alpha = 0.1) +
  coord_cartesian(xlim = c(min(foxes$area), max(foxes$area)), ylim = c(min(foxes$weight), max(foxes$weight)))

p_foxes %>% 
  mutate(intercept = alpha + beta1 * mean(foxes$area)) %>% 
  ggplot(aes(x = groupsize, y = weight)) +
  geom_abline(aes(intercept = intercept, slope = beta2), alpha = 0.1) +
  coord_cartesian(xlim = c(min(foxes$groupsize), max(foxes$groupsize)), ylim = c(min(foxes$weight), max(foxes$weight)))
```

This suggests groupsize is masking area. That is, groupsize is correlated with area, and they are both correlated with weight but have opposite signs. Groupsize is negatively correlated with weight and area is positively correlated with weight. 

```{r}
m_h3_area <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta_a * area + beta_g * groupsize,
    alpha ~ dnorm(4, 10),
    beta_a ~ dnorm(0, 3),
    beta_g ~ dnorm(0, 3),
    sigma ~ dexp(1)
  ),
  data = foxes
)

m_h3_food <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta_g * groupsize + beta_f * avgfood,
    alpha ~ dnorm(4, 10),
    beta_g ~ dnorm(0, 3),
    beta_f ~ dnorm(0, 3),
    sigma ~ dexp(1)
  ),
  data = foxes
)

m_h3 <- quap(
  flist = alist(
    weight ~ dnorm(mu, sigma),
    mu <- alpha + beta_a * area + beta_g * groupsize + beta_f * avgfood,
    alpha ~ dnorm(4, 10),
    beta_a ~ dnorm(0, 3),
    beta_g ~ dnorm(0, 3),
    beta_f ~ dnorm(0, 3),
    sigma ~ dexp(1)
  ),
  data = foxes
)
precis(m_h3_area)
precis(m_h3_food)
precis(m_h3)
```

I would include area before food, since its estimate has a lower variance. Since area and food are correlated, including both increases the variance of the estimate for both.
